# R Shiny Here We Go

I'm starting a new job in a few weeks and I need to get up to speed with R Shiny. Here goes.

```{r, include = FALSE}
# Libraries
library(glue)
library(janitor)
library(knitr)
library(readxl)
library(tidyverse)
library(utils)
library(dtplyr)
library(openxlsx)

# Parameters
url <- "http://open-data-assurance-maladie.ameli.fr/
            depenses/download_file.php?file=Open_DAMIR/A201912.csv.gz"
filename <- str_split(url, "/")[[1]] %>% last()
filename_root <- str_split(filename, "\\.")[[1]][1]
filename_extension <- str_split(filename, "\\.")[[1]][2]
filename_compression <- str_split(filename, "\\.")[[1]][3]

# Seed
set.seed(451)

#################################################
# FUNCTIONS
#################################################

## ASSIGN_FACTOR_LABELS

### 1) Transforms the column in df to a factor
### 2) Retrieves the factor labels from Lexique_open-DAMIR
### ! There are NAs in the PRS_NAT sheet, hence the drop_na()
assign_factor_labels <- function(df, column) {
    df %>%
    mutate(
        across(
            column,
            ~ .x %>% factor(
                levels = read_xls(
                    "../references/Lexique_open-DAMIR.xls",
                    sheet = column
                )[1] %>%
                drop_na() %>%
                as_vector(),
                labels = read_xls(
                    "../references/Lexique_open-DAMIR.xls",
                    sheet = column
                )[2] %>%
                drop_na() %>%
                modify(str_to_sentence) %>%
                as_vector()
            )
        )
    )
}
```

## Downloading stuff

There may be a Captcha. If there is indeed one, then download the file by hand. Sorry!

The extraction takes forever. Do it by hand; programming this was actually a tremendous waste of my time.

```{r, include = FALSE}
# Download file
# Unless it is already there, ofc
if (file.exists(glue("../data/raw/{filename}"))) {
    glue("File {filename} already exists")
} else {
    download.file(
        url = url,
        destfile = glue("../data/raw/{filename}")
    )
}

# Extract
if (file.exists(glue("../data/raw/{filename_root}.{filename_extension}"))) {
    glue("File {filename_root}.{filename_extension} already exists")
} else {
    to_extract <- readr::read_csv(glue("../data/raw/{filename}"))
    to_extract %>% readr::write_csv(
        glue("../data/raw/{filename_root}.{filename_extension}")
    )
    rm(to_extract)
    gc()
}
```

This file is way too big: 35M rows and 56 variables. I'll take a random sample of 30K of them.

```{r, include = FALSE}
if (file.exists(glue("../data/raw/{filename_root}_sample.csv"))) {
    print("There is already a sample file.")
} else {
    df <- readr::read_csv2(
        glue("../data/raw/{filename_root}.{filename_extension}")
    )
    df %>%
        slice_sample(n = 30000) %>%
        readr::write_csv(
            glue("../data/raw/{filename_root}_sample.csv")
        )
}

```

```{r, include = FALSE}
# Importing df
df <- readr::read_csv(
        glue("../data/raw/{filename_root}_sample.csv")
    ) %>%
    select(-`...56`) # Why is this there?


# Importing labels
df_labels <-
    read_xls(
        "../references/Lexique_open-DAMIR.xls",
        skip = 2
    ) %>%
    select(Variable...1, Libellé...2) %>%
    rename(
        variable = Variable...1,
        label = Libellé...2
    ) %>%
    slice(-1)


# List of factors in the df
factors_list <- excel_sheets("../references/Lexique_open-DAMIR.xls") %>%
    keep(~ any(colnames(df) %in% .x))

# Transform categorical variables in the df to factors
# Factor labels are retrieved from Lexique_open-DAMIR
## I am ashamed of using a for loop, but I haven't found a way to elegantly
## put this function in a mutate(across())
for (var in factors_list) {
    print(glue("Assigning values of factor {var}"))
    df <- df %>%
        assign_factor_labels(var)
}

# Add variable class to labels dataframe
df_labels <- df_labels %>%
    left_join(
        bind_cols(
            variable = df %>% colnames(),
            class = df %>% sapply(class)
        )
    )

saveRDS(df, "../data/final/A201912.RDS")
saveRDS(df_labels, "../data/final/A201912_labels.RDS")
```

```{r}
df
df_labels %>% kable()
```


# One-way tables templates

```{r}
df %>%
    select_if(is.factor) %>%
    select(-PRS_NAT) %>%
    sapply(tabyl) %>%
    adorn_pct_formatting(digits = 2)
```

```{r}
tabyl(df$AGE_BEN_SNDS) %>%
    adorn_pct_formatting(digits = 2)
```

```{r}
    df %>%
    summarise(
        N = n(),
        Mean = mean(FLT_ACT_COG),
        SD = sd(FLT_ACT_COG),
        Minimum = min(FLT_ACT_COG),
        `1%` = quantile(FLT_ACT_COG, 0.01),
        `5%` = quantile(FLT_ACT_COG, 0.05),
        `25%` = quantile(FLT_ACT_COG, 0.25),
        `50%` = quantile(FLT_ACT_COG, 0.50),
        `75%` = quantile(FLT_ACT_COG, 0.75),
        `95%` = quantile(FLT_ACT_COG, 0.95),
        `99%` = quantile(FLT_ACT_COG, 0.99),
        Maximum = max(FLT_ACT_COG)
    )
```

# Two-way tables templates

```{r}
varlist <- c("AGE_BEN_SNDS", "BEN_SEX_COD")
for (var in varlist) {
    table <- df %>%
        tabyl(!!sym(var))
    print(table)
}
```

```{r}
df %>%
    tabyl(AGE_BEN_SNDS, BEN_SEX_COD) %>%
    adorn_totals() %>%
    adorn_percentages("row") %>%
    adorn_pct_formatting(digits = 2) %>%
    adorn_ns() %>%
    as.tibble()
```

```{r}
df_labels %>%
    filter(variable == "AGE_BEN_SNDS") %>% 
    pull(label)
```

```{r}
df %>% 
    filter(AGE_BEN_SNDS %in% c("70 - 79 ans")) %>%
    mutate(AGE_BEN_SNDS = fct_drop(AGE_BEN_SNDS)) %>%
    pull(AGE_BEN_SNDS) %>%
    is.factor()
```

```{r}
bind_rows(
    df %>%
        tabyl(AGE_BEN_SNDS) %>%
        adorn_totals() %>%
        slice_tail(),
    df %>%
        tabyl(AGE_BEN_SNDS)
)
```

```{r}
df %>%
    group_by(BEN_SEX_COD) %>%
    summarise(
        N = n(),
        Mean = mean(PRS_REM_MNT),
        SD = sd(PRS_REM_MNT),
        Minimum = min(PRS_REM_MNT),
        `1%` = quantile(PRS_REM_MNT, 0.01),
        `5%` = quantile(PRS_REM_MNT, 0.05),
        `25%` = quantile(PRS_REM_MNT, 0.25),
        `50%` = quantile(PRS_REM_MNT, 0.50),
        `75%` = quantile(PRS_REM_MNT, 0.75),
        `95%` = quantile(PRS_REM_MNT, 0.95),
        `99%` = quantile(PRS_REM_MNT, 0.99),
        Maximum = max(PRS_REM_MNT)
    )
```

```{r}
df %>% drop_na(PRS_REM_MNT, BEN_SEX_COD)
```

```{r}
df %>%
    tabyl(BEN_SEX_COD)
```

# Custom report dev

```{r}

# SETUP & STYLES
wb <- createWorkbook("Sample Report")
addWorksheet(wb, "Table 1")

hs1 <- createStyle(
    border = "Top"
)
bs1 <- createStyle(
    border = "Bottom"
)
indent_style <- createStyle(
    indent = 1
)
rounding = 2

# Parameters
colvar <- "BEN_SEX_COD"
rowvars <- c("AGE_BEN_SNDS", "ATT_NAT")
percentage_style <- "row"
total_col <- TRUE
pval_col <- FALSE

# Counters
row_counter <- 1

# Column variable categories
colvar_categories <- df %>% pull(colvar) %>% levels()



###############################################################################
# HEADER

# Build header
table_header <- df %>%
    tabyl(!!sym(rowvars[1]), !!sym(colvar)) %>%
    as.tibble() %>%
    slice(0)
# Total column
if (total_col) {
    table_header <- table_header %>%
    mutate(Total = NA)
}
## % (N) in header
header_legend <- rep(c("% (N)"), times = length(table_header)-1) %>%
    t()
# P-value in a column
if (pval_col) {
    table_header <- table_header %>%
    mutate(`P-value` = NA)
}
## N & % of column variable
colvar_count <- df %>%
    tabyl(BEN_SEX_COD) %>%
    {if (total_col) adorn_totals(.) else .} %>%
    adorn_pct_formatting(2) %>%
    mutate(n = paste0("(", as.character(n), ")")) %>%
    unite("percent", percent:n, sep = (" ")) %>%
    mutate(percent = str_remove(percent, "%")) %>%
    pull(percent) %>%
    as_tibble() %>%
    rename(Total = value) %>%
    t()

## Write colvar name
writeData(wb = wb, sheet = "Table 1", x = colvar, startCol = 2)
addStyle(wb, "Table 1", cols = 1:length(table_header), rows = 1, style = hs1)
mergeCells(wb, "Table 1", cols = (1:length(colvar_count)) + 1, rows = 1)
row_counter <- row_counter + 1
## Write colvar categories
writeData(wb = wb, sheet = "Table 1", x = table_header, startRow = row_counter)
## Remove variable name in header
writeData(wb = wb, sheet = "Table 1", x = " ", startRow = row_counter)
row_counter <- row_counter + 1
## Add header legend: % (N)
writeData(wb = wb, sheet = "Table 1", x = header_legend, startCol = 2,
    startRow = row_counter, colNames = FALSE)
addStyle(wb, "Table 1", cols = 1:length(table_header), style = bs1,
    rows = row_counter)
row_counter <- row_counter + 1
## Add counts of col var
writeData(wb = wb, sheet = "Table 1", x = colvar_count, startCol = 1,
    startRow = row_counter, colNames = FALSE, rowNames = TRUE)
row_counter <- row_counter + 1

###############################################################################
# TABLE CONTENT
for (rowvar in rowvars) {
    # Create table
    table <- df %>%
        tabyl(!!sym(rowvar), !!sym(colvar)) %>%
        adorn_percentages(percentage_style) %>%
        adorn_pct_formatting(digits = rounding) %>%
        mutate(across(colvar_categories, ~ str_remove(.x, "%"))) %>%
        as.tibble()

    # Total column
    if (total_col) {
        table_col_total <- df %>%
            tabyl(!!sym(rowvar)) %>%
            adorn_pct_formatting(2) %>%
            mutate(n = paste0("(", as.character(n), ")")) %>%
            unite("percent", percent:n, sep = (" ")) %>%
            mutate(percent = str_remove(percent, "%")) %>%
            rename(Total = percent)

        table <- table %>%
            left_join(table_col_total)
    }

    # Table
    ## Write variable name
    writeData(wb = wb, sheet = "Table 1", x = rowvar,
        startRow = row_counter)
    row_counter <- row_counter + 1
    ## Write content
    writeData(wb, "Table 1", table, startRow = row_counter, colNames = FALSE)
    ## Indent rowvar categories
    addStyle(wb, "Table 1", cols = 1,
        rows = row_counter:(row_counter + nrow(table) - 1), style = indent_style)
    row_counter <- row_counter + nrow(table)
}

###############################################################################
# FOOTER

# Border to bottom of table
addStyle(wb, "Table 1", cols = 1:length(table_header),
    rows = row_counter-1, style = bs1, stack = TRUE)

###############################################################################
# WRITE TABLE
openXL(wb)
saveWorkbook(wb, "Test.xlsx", overwrite = TRUE)
```

```{r}
# Functions

ds_report <- function(data, workbook, worksheet, colvar, rowvars, rounding = 2,
    percentage_style = "row", counts = FALSE,
    separate_counts = FALSE, total_col = TRUE, pval_col = FALSE,
    chi_cat_reject = FALSE, chi_cat_reject_threshold = 30, filename) {

# Data: dataframe
# workbook: name of the workbook (string)
# worksheet: name of the worksheet (string)
# colvar: name of column variable (string, max one)
# rowvars: names of row variables (vector of strings)
# rounding: number of digits for rounding
# percentage_style: "row", "col" or "both" (= cell) percentages
# counts: show counts in parenthesis
# separate_counts: show counts in a new column
# total_col: add a total column (bool)
# pval_col: put the p-values in a new column instead of new rows (bool)
# chi_cat_reject: Automatic drop of categories with low counts
# chi_cat_reject_threshold: Threshold at which categories are dropped (default:  categories with n <= 30 are dropped)
## Use a value >= to 1 to drop based on counts (n), or < 1 to drop based on percentages (e.g. 0.01)
# filename: Name of excel file to export to

    # SETUP & STYLES
    wb <- createWorkbook(workbook)
    addWorksheet(wb, worksheet)

    hs1 <- createStyle(
        border = "Top"
    )
    bs1 <- createStyle(
        border = "Bottom"
    )
    indent_style <- createStyle(
        indent = 1
    )

    # Counters
    row_counter <- 1

    # Column variable categories
    colvar_categories <- data %>%
        pull(colvar) %>%
        levels()

    # Subfunction
    separate_results <- function(data, column) {
        column_n <- paste0(column, "_N")
        column_perc <- paste0(column, "_N")

        data %>%
            mutate(!!sym(column) := str_remove(!!sym(column), "\\)")) %>%
            separate(
                !!sym(column),
                c(paste0(column, "_N"), paste0(column, "_%")),
                sep = "\\("
            ) %>%
            mutate(!!sym(column_n) := str_trim(!!sym(column_n)))
    }

    ###########################################################################
    # HEADER

    # Build header
    table_header <- data %>%
        tabyl(!!sym(rowvars[1]), !!sym(colvar)) %>%
        as.tibble() %>%
        slice(0)
    # Total column
    if (total_col) {
        table_header <- table_header %>%
        mutate(Total = NA)
    }
    ## % (N) in header
    if (counts & separate_counts) {
        header_legend <- rep(c("%", "N"), times = length(table_header)-1) %>%
            t()
    } else {
        header_legend <- rep(c("% (N)"), times = length(table_header)-1) %>%
            t()
    }
    
    # P-value in a column
    if (pval_col) {
        table_header <- table_header %>%
        mutate(`P-value` = NA)
    }
    ## COLUMN VARIABLE TREATMENT
    colvar_count <- data %>%
        tabyl(!!sym(colvar))
    ### FOR CHI-SQUARED TESTS
    ### Used to filter some columns in the chi-squared tests
    if (chi_cat_reject) {
        colvar_cats_chi_reject <- colvar_count %>%
        {if (chi_cat_reject_threshold >= 1)
            filter(., n < chi_cat_reject_threshold) else
            filter(., percent < chi_cat_reject_threshold)} %>%
        pull(BEN_SEX_COD) %>%
        as.vector()
        message("Rejected column categories for chi-sq test: ", colvar_cats_chi_reject)
    }
    ### Get counts (N + %) of column variable
    colvar_count <- colvar_count %>%
        {if (total_col) adorn_totals(.) else .} %>%
        adorn_pct_formatting(2) %>%
        mutate(n = paste0("(", as.character(n), ")")) %>%
        unite("percent", percent:n, sep = (" ")) %>%
        mutate(percent = str_remove(percent, "%")) %>%
        pull(percent) %>%
        as_tibble() %>%
        rename(Total = value) %>%
        t()

    if (counts & separate_counts) {
        colvar_count <- colvar_count %>% as_tibble()
        for (var in colnames(colvar_count)) {
            colvar_count <- colvar_count %>% separate_results(var)
        }
        colvar_count <- colvar_count %>%
            mutate(
                rowname = "Total"
            ) %>%
            column_to_rownames("rowname")
    }

    ## Write colvar name
    writeData(wb = wb, sheet = worksheet, x = colvar, startCol = 2)
    addStyle(wb, worksheet, cols = 1:(length(colvar_count)+1),
        rows = 1, style = hs1
    )
    mergeCells(wb, worksheet, cols = (1:length(colvar_count)) + 1, rows = 1)
    row_counter <- row_counter + 1
    ## Write colvar categories
    writeData(wb = wb, sheet = worksheet, x = table_header,
        startRow = row_counter
    )
    ## Remove variable name in header
    writeData(wb = wb, sheet = worksheet, x = " ", startRow = row_counter)
    
    ### Alternative header if N and % are separated
    if (counts & separate_counts) {
        alt_header_column = 2
        for (cat in colvar_categories) {
            writeData(wb = wb, sheet = worksheet, x = cat, 
                startRow = row_counter, startCol = alt_header_column)
            mergeCells(wb, worksheet, cols = (alt_header_column:(alt_header_column+1)),
                rows = row_counter)
            alt_header_column <- alt_header_column + 2
        }
        if (total_col) {
            writeData(wb = wb, sheet = worksheet, x = "Total", 
                startRow = row_counter, startCol = alt_header_column)
            mergeCells(wb, worksheet, cols = (alt_header_column:(alt_header_column+1)),
                rows = row_counter)
            alt_header_column <- alt_header_column + 2
        }
    }

    row_counter <- row_counter + 1
    ## Add header legend: % (N)
    writeData(wb = wb, sheet = worksheet, x = header_legend, startCol = 2,
        startRow = row_counter, colNames = FALSE)
    addStyle(wb, worksheet, cols = 1:(length(colvar_count)+1), style = bs1,
        rows = row_counter)
    row_counter <- row_counter + 1
    ## Add counts of col var
    writeData(wb = wb, sheet = worksheet, x = colvar_count, startCol = 1,
        startRow = row_counter, colNames = FALSE, rowNames = TRUE)
    row_counter <- row_counter + 1

    ###########################################################################
    # TABLE CONTENT
    for (rowvar in rowvars) {
        # Create table
        table <- data %>%
            tabyl(!!sym(rowvar), !!sym(colvar)) %>%
            adorn_percentages(percentage_style) %>%
            adorn_pct_formatting(digits = rounding) %>%
            {if (counts) adorn_ns(.) else .} %>%
            mutate(across(colvar_categories, ~ str_remove(.x, "%"))) %>%
            as.tibble()

        # Total column
        if (total_col) {
            table_col_total <- data %>%
                tabyl(!!sym(rowvar)) %>%
                adorn_pct_formatting(2) %>%
                mutate(n = paste0("(", as.character(n), ")")) %>%
                unite("percent", percent:n, sep = (" ")) %>%
                mutate(percent = str_remove(percent, "%")) %>%
                rename(Total = percent)

            table <- table %>%
                left_join(table_col_total)

        }
        
        # Separate counts & percentages
        if (counts & separate_counts) {
            for (var in colvar_categories) {
                table <- table %>% separate_results(var)
            }
            if (total_col) {
                table <- table %>% separate_results("Total")
            }
        }

        # Calculate chisq test
        table_chi <- data %>%
            mutate(
                # Tabyl shows unused categories
                # These should not be used in the chisq test
                !!sym(rowvar) := fct_drop(!!sym(rowvar))
            ) %>%
            tabyl(!!sym(rowvar), !!sym(colvar)) %>%
            select(-colvar_cats_chi_reject, -!!sym(rowvar)) %>%
            as_tibble()

        chi_pval <- if (chisq.test(table_chi)$p.value < 0.001) {
            "< 0.001"
        } else if (chisq.test(table_chi)$p.value < 0.05) {
            chisq.test(table_chi)$p.value %>% round(3)
        } else {
            chisq.test(table_chi)$p.value %>% round(2)
        }

        # Table
        ## Write variable name
        writeData(wb = wb, sheet = worksheet, x = rowvar,
            startRow = row_counter
        )
        row_counter <- row_counter + 1
        ## Write content
        writeData(wb, worksheet, table, startRow = row_counter,
            colNames = FALSE
        )
        ## Indent rowvar categories
        addStyle(wb, worksheet, cols = 1,
            rows = row_counter:(row_counter + nrow(table) - 1),
            style = indent_style
        )
        row_counter <- row_counter + nrow(table)
        ## Add p-value
        writeData(wb, worksheet, "P-value", startRow = row_counter)
        writeData(wb, worksheet, chi_pval, startRow = row_counter, startCol = 2)
        addStyle(wb, worksheet, cols = 1,
            rows = row_counter,
            style = indent_style
        )
        row_counter <- row_counter + 1
    }

    ###########################################################################
    # FOOTER

    # Border to bottom of table
    addStyle(wb, worksheet, cols = 1:(length(colvar_count)+1),
        rows = row_counter-1, style = bs1, stack = TRUE)

    ###########################################################################
    # WRITE TABLE
    openXL(wb)
    saveWorkbook(wb, filename, overwrite = TRUE)
}
```

```{r}
# df_report <- df %>%
#     filter(BEN_SEX_COD %in% c("Masculin", "Feminin")) %>%
#     mutate(BEN_SEX_COD = fct_drop(BEN_SEX_COD))

ds_report(
    data = df_report,
    workbook = "Sample report",
    worksheet = "Table 1", 
    colvar = "BEN_SEX_COD",
    rowvars = c("AGE_BEN_SNDS", "ATT_NAT", "PRS_PDS_QCP"),
    counts = FALSE,
    separate_counts = FALSE,
    total_col = TRUE,
    pval_col = FALSE,
    chi_cat_reject = TRUE,
    chi_cat_reject_threshold = 30,
    filename = "Test.xlsx"
)
```

```{r}
colvar_cats_chi_reject <- c("Inconnu ou personne morale sans civilite", "Valeur inconnue")
rowvar <- "ATT_NAT"
colvar <- "BEN_SEX_COD"
total_col <- TRU
percentage_style <- "row"
rounding <- 2
counts <- TRUE
separate_counts <- TRUE


colvar_count <- df %>%
        tabyl(!!sym(colvar))
colvar_count

colvar_count %>%
    {if (total_col) adorn_totals(.) else .} %>%
    adorn_pct_formatting(2) %>%
    mutate(percent = str_remove(percent, "%")) %>%
    select(-!!sym(colvar))


colvar_count <- colvar_count %>%
        {if (total_col) adorn_totals(.) else .} %>%
        adorn_pct_formatting(2) %>%
        mutate(n = paste0("(", as.character(n), ")")) %>%
        unite("percent", percent:n, sep = (" ")) %>%
        mutate(percent = str_remove(percent, "%")) %>%
        pull(percent) %>%
        as_tibble() %>%
        rename(Total = value) %>%
        t()

# SEPARATE TOTAL N AND %
colvar_count <- colvar_count %>% as_tibble()
for (var in colnames(colvar_count)) {
    colvar_count <- colvar_count %>% separate_results(var)
}
colvar_count %>% mutate(
    rowname = "Total"
) %>% column_to_rownames("rowname")

colvar_categories <- df %>%
    pull(colvar) %>%
    levels()

table <- df %>%
            tabyl(!!sym(rowvar), !!sym(colvar)) %>%
            adorn_percentages(percentage_style) %>%
            adorn_pct_formatting(digits = rounding) %>%
            {if (counts) adorn_ns(.) else .} %>%
            mutate(across(colvar_categories, ~ str_remove(.x, "%"))) %>%
            as.tibble()

        # Total column
        if (total_col) {
            table_col_total <- df %>%
                tabyl(!!sym(rowvar)) %>%
                adorn_pct_formatting(2) %>%
                mutate(n = paste0("(", as.character(n), ")")) %>%
                unite("percent", percent:n, sep = (" ")) %>%
                mutate(percent = str_remove(percent, "%")) %>%
                rename(Total = percent)

            table <- table %>%
                left_join(table_col_total)


table_header <- df %>%
        tabyl(!!sym(rowvar), !!sym(colvar)) %>%
        as.tibble() %>%
        slice(0)

separate_results <- function(data, column) {
    column_n <- paste0(column, "_N")
    column_perc <- paste0(column, "_N")

    data %>%
        mutate(!!sym(column) := str_remove(!!sym(column), "\\)")) %>%
        separate(
            !!sym(column),
            c(paste0(column, "_N"), paste0(column, "_%")),
            sep = "\\("
        ) %>%
        mutate(!!sym(column_n) := str_trim(!!sym(column_n)))
}

for (var in colvar_categories) {
    table <- table %>% separate_results(var)
}

```